%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{edXchapter}{Advanced Examples}

\begin{edXsection}{Matrices}

\begin{edXvertical}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{edXproblem}{Matrix expression equality}

This problem demonstrates testing of matrix expression equality.  It properly handles the non-commutativity of matrices.

Let $X$, $Y$, and $Z$ be the usual Pauli matrices,
\bea
     	X &=& \mattwoc{0}{1}{1}{0}
\\ 	Y &=& \mattwoc{0}{-i}{i}{0}
\\ 	Z &=& \mattwoc{1}{0}{0}{-1}
\,.
\eea
Give an expression $M$ equal to $iZ$.

Valid answers include {\tt\LARGE X*Y} and {\tt\LARGE i*Z}.  But {\tt\LARGE Y*X} is properly graded as being invalid.

Note that {\LARGE\tt  X * exp(-i*Y) * Y * exp(i*Y)} also works (matrix
exponentials do the correct thing -- although, unfortunately, the edX
edge system doesn't provide {\LARGE\tt scipy.linalg}, and so it will
not work there).

% The following code imports matrix_evaluator from the course
% /python_lib directory of python code.
%
% Using the library makes it possible to efficiently re-use code
% and write modular code with well-defined scope.
%
% This works fine with an LMS using direct import of XML code.
%
% Unfortunately, the edX Studio system is broken, in that it does
% not allow use of /python_lib or /code directories of python
% code libraries.  Thus, for this demo to work on the edx.org
% Studio system, we are forced to paste all the python code
% here, instead of importing it using normal semantics.
%
% \begin{edXscript}
% 
% from matrix_evaluator import *
% 
% \end{edXscript}

% This is the line which pastes in all the python code,
% needed to work on edx.org
\edXincludepy{../python_lib/all_evaluators_in_one.py}

%  options="samples='X,Y,Z@[0|1;1|0],[0|(0-1j);(0+1j)|0],[1|0;0|-1]:[0|1;1|0],[0|(0-1j);(0+1j)|0]],[1|0;0|-1]j#50'!altanswer='-Y*X'!altanswer='i*Z'"

\edXinline{$M = $ }
\edXabox{expect="X*Y" 
  type="custom" cfn="test_formula" 
  inline='1'
  math="1" 
  size="70"
  options="samples='X,Y,Z@[0|1;1|0],[0|(0-1j);(0+1j)|0],[1|0;0|-1]:[0|1;1|0],[0|(0-1j);(0+1j)|0]],[1|0;0|-1]#50'"
  preprocessorClassName="MathjaxPreprocessorForQM" preprocessorSrc="/static/js/mathjax_preprocessor_for_QM.js"
 }%

  % options="samples='X,Y,i@[1|2;3|4],[0|2;4|6],0+1j:[5|5;5|5],[8|8;8|8],0+1j#50'!altanswer='-Y*X'!altanswer='i*Z'"

\begin{edXshowhide}{studioproblem}{Implementation notes and limits of edX studio}
The checking of this formula is done using a somewhat extensive python
script.  The edX platform allows such python scripts to be stored in
the course /python_lib or /code directory.  This means that each
problem can simply have a short script with a python ``import'' line
to retrieve the library functions.

This works very well when you are authoring the course directly in
XML, and not via uploading to edX Studio.

Unfortunately, the edX Studio system is broken, in that it does
not allow use of /python_lib or /code directories of python
code libraries.  Thus, for this demo to work on the edx.org
Studio system, we are forced to paste all the python code
in each problem, instead of importing it using normal semantics.

Note that edX Studio also does not allow a hierarchy of directories in
the /static directory: all files uploaded to Studio are stored flat in
that directory.  This means that HTML or javascript files which
attempt to load files from subdirectories will break, when using
Studio.  

Alas.  Maybe someday Studio will be fixed to work better.  
\end{edXshowhide}

\begin{edXsolution}

Either $XY$ or $-YX$ but not $YX$.

\end{edXsolution}

\edXgitlink{\giturl}{Source TeX}
\end{edXproblem}

\end{edXvertical}

\begin{edXvertical}

\begin{edXproblem}{Quantum mechanics ket input}{attempts=1000 weight=10}

Quantum mechanics uses ``ket'' notation like $|\psi\>$ to indicate unit
vectors.

This problem demonstrates how to use customresponse python code to
check for equality of formulas with such ket notation, as well as a
javascript pre-processor to properly format ket expressions.

Type in {\tt (2*|0>+3*|b>)/sqrt(13)} as the answer below.

Multiple correct answer are supported.  {\tt |b>} is also correct, in
this example.

\edXxml{<font color="red"><b> Remeber to explicitly denote multiplication using [mathjaxinline]\tt *[/mathjaxinline]</b></font>}

\edXbr

% The following code imports matrix_evaluator from the course
% /python_lib directory of python code.
%
% Using the library makes it possible to efficiently re-use code
% and write modular code with well-defined scope.
%
% This works fine with an LMS using direct import of XML code.
%
% Unfortunately, the edX Studio system is broken, in that it does
% not allow use of /python_lib or /code directories of python
% code libraries.  Thus, for this demo to work on the edx.org
% Studio system, we are forced to paste all the python code
% here, instead of importing it using normal semantics.
%
% \begin{edXscript}
% 
% from ket_evaluator import *
% 
% \end{edXscript}

% This is the line which pastes in all the python code,
% needed to work on edx.org
\edXincludepy{../python_lib/all_evaluators_in_one.py}

% Note the samples line below,
%
%  options="samples='a,b@1,1:20,20#10'!altanswer='|a>'"
%
% This specifies only 10 random samples (the number after #) 
% whereas to be more certain, one would probably want
% more samples.  But the edx.org system limits
% cpu and memory resources of the checking script, and
% throws an error if the number of samples for ket
% equality is greater than around this number.
%
% On the residential MITx system the number of
% samples used in this example can be 50 (and maybe higher).

\edXinline{$|\lambda\> = $ }
\edXabox{expect="(2*|0>+3*|b>)/sqrt(13)" 
  type="custom" cfn="test_ket" inline='1'
  math="1" 
  size="60"
  options="samples='a,b@1,1:20,20#10'!altanswer='|a>'"
  preprocessorClassName="MathjaxPreprocessorForQM" preprocessorSrc="/static/js/mathjax_preprocessor_for_QM.js"
 }%

Note that correctness of the student's input expression is checked by
python code running on the server.  edx.org limits CPU time and memory
available to the code, and thus only 10 random vector samples can be
checked on this site.  Other systems, such as the residential MITx
system, are more accommodating, and allow enough resources to check 50
or more random samples, thus lowering the probability of mistaken
correctness.

\begin{edXsolution}

The $\hat{n}$ vector goes through the point $(\theta, \phi)$ specified
in polar coordinates, with $\phi=0$ and 
\bea
   \theta = \cos^{-1} \frac{\omega_0/2}{\sqrt{V^2 + \left(\frac{\omega_0}{2}\right)^2}} 
   = \tan^{-1}\left(\frac{2V}{\omega_0}\right)
\,.
\eea
In terms $\theta$, the eigenstates may be expressed as
\bea
	|\lambda_+\> = \cos\frac{\theta}{2} |a\> +\sin\frac{\theta}{2} |b\>
\\
	|\lambda_-\> = -\sin\frac{\theta}{2} |a\> +\cos\frac{\theta}{2} |b\>
\,.
\eea

\end{edXsolution}


\end{edXvertical}

\end{edXsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{edXsection}{Adaptive hints}

\begin{edXvertical}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{edXproblem}{Magnitude off hint}{attempts=500}

\begin{edXscript}
eh = {'range':[ 280, 290 ],
      'hint': "You are assuming too low a temperature for the hot dry rock"
     }
ans0 = 177
expected=[ {'val': ans0, 'extra_hints': [eh]} ]
\end{edXscript}

\edXincludepy{XML/magnitude_hint2.py}

This problem demonstrates the ability to give hints when students
enter numerical answers which are too far from the expected result.

The correct answer is 177.  Try typing in 1 or 280 instead (you'll get hints).

How does the \href{http://en.wikipedia.org/wiki/Energy_density}{energy
  density} of \href{http://en.wikipedia.org/wiki/Petroleum}{crude oil}
$E_{\rm oil}$ compare with the energy density of hot
\href{http://en.wikipedia.org/wiki/Geothermal_energy}{geothermal}
fluid $E_{\rm geo}$?

Model the geothermal fluid as water, and assume its starts at 360 K
when pumped from oil/geothermal wells at a rate of 30 l/s, and is
cooled to 300 K.

Give the ratio of energy densities.

\edXinline{$E_{\rm oil}/E_{\rm geo}$ =} \edXabox{expect="\$ans0" type="numerical"
  tolerance='20\%' inline='1' hintfn="hint_mag0" } 
 
\begin{edXsolution}

  If the geothermal fluid is basically just water, then its density at
  360~K is approximately 1~g/cm$^3$ (really is 0.97~g/cm$^3$ but also
  changes with temperature as heat is extracted). The heat capacity is
  approximately $c = 4.18~kJ/kg K$. If the fluid is cooled to a
  typical temperature of 300~K, then the total energy density of the
  geothermal fluid is
\begin{equation}
\mathcal{E}_{\rm geo} = \rm(4.18~kJ/kg~K)(60~K) = 251~kJ/kg
\end{equation}
Oil has an energy density of around 43~MJ/kg (wikipedia gives 46
MJ/kg).  Thus, $E_{\rm oil}/E_{\rm geo} \approx 177$, so oil has
nearly 200 times the energy density of the geothermal fluid.  

\end{edXsolution}

\edXgitlink{\giturl}{Source TeX}
\end{edXproblem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{edXvertical}

\end{edXsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{edXsection}{Drag and drop input}

\begin{edXvertical}

\begin{edXproblem}{Approximate energy level diagram}{attempts=10}

Use your knowledge of the size of the Lamb shift, and the scale of fine
structure splittings, to properly label the following energy level diagram.

\edXinclude{XML/ps4-hydrogen-dnd.xml}

\begin{edXsolution}

\begin{itemize}

\item The fine structure splitting raises the four $2P_{3/2}$ states about
10 GHz above the two $2P_{1/2}$ and two $2S_{1/2}$ states.

\item The Lamb shift raises $2S_{1/2}$ above $2P_{1/2}$ by 1.06 GHz.

\end{itemize}

\end{edXsolution}

\edXgitlink{\giturl}{Source TeX}
\end{edXproblem}


\end{edXvertical}

%%%%%%%%%%%%%%%%%%%%

\begin{edXvertical}

\begin{edXproblem}{Drag and drop problem created via latex2dnd}{attempts=1000}

The \href{https://github.com/mitocw/latex2dnd}{latex2dnd} package can
be used to generate edX
\href{http://edx.readthedocs.org/projects/devdata/en/latest/course_data_formats/drag_and_drop/drag_and_drop_input.html}{drag-and-drop}
problems.  These problems are very convenient for complex equations,
or for labeling of graphical images:

\begin{edXshowhide}{dndex}{LaTeX code for quadratic polynomial example}
Download \href{https://github.com/mitocw/latex2dnd/blob/master/latex2dnd/testtex/quadratic.tex}{from github}:
\begin{verbatim}
\documentclass{article}
\input{latex2dnd}

\begin{document}

% define drag-drop labels

\DDlabel{term1}{$-b$}
\DDlabel{term1p}{$+b$}
\DDlabel{term2}{$b^2$}
\DDlabel{dubexp}{$b^{2^\alpha}$}
\DDlabel{dubsub}{$b_{2_\alpha}$}
\DDlabel{fac}{$-4ac$}
\DDlabel{facp}{$+4ac$}
\DDlabel{ta}{$+2a$}
\DDlabel{tam}{$-2a$}

% shorthand macro to make all boxes the same size (6 by 4)
\newcommand\DDB[2]{\DDbox{#1}{6ex}{4ex}{#2}}

% the formula with boxes
$$\lambda = \frac{\DDB{1}{term1}\pm \sqrt{\DDB{2}{term2}\DDB{3}{fac}}}{\DDB{4}{ta}}$$

% output labels, with fixed box heights
\writeDDlabels[4.3ex]

\end{document}
\end{verbatim}
\end{edXshowhide}

Complete this to give the correct equation for the roots of a
quadratic polynomial $ax^2 + bx + c$:

\edXinclude{dnd/quadratic_dnd.xml}

\edXgitlink{\giturl}{Source TeX}
\end{edXproblem}

%%%%%%%%%%

\end{edXvertical}

\end{edXsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{edXsection}{Graphical input using javascript}

\begin{edXvertical}

\begin{edXproblem}{Graphical input via jsxgraph}{attempts=1000}

The \href{http://edx.readthedocs.org/projects/devdata/en/latest/course_data_formats/jsinput.html}{jsinput} framework provides a very convenient way to
employ virtually arbitrary javascript code as a user interface for
input into edX auto-graded problems.  

This is an example which uses sliders and graphs from
\href{http://jsxgraph.uni-bayreuth.de/}{jsxgraph}, a nice javascript
math function plotting package, as input to a physics problem.

     Identify one such field-independent transition in the figure
     below, by doing the following:
     \begin{itemize}
     \item Use the slider to select a value of $x$
     \item Choose two different states (A...H) for the transition
     \item Click ``Check'' to check your answer
     \end{itemize}

\edXincludepy{XML/test_findep1.py}

% this javascript forces the saved jsinput state to be reloaded after 1.4 seconds
\edXxml{<script type="text/javascript">setTimeout(function(){ JSInput.jsinputConstructor(\$('section.jsinput'))}, 1400);</script>}

\edXabox{expect="" type="jsinput" cfn="test_findep" 
  width="650"
  height="555"
  gradefn="getinput"
  get_statefn="getstate"
  set_statefn="setstate"
  html_file="/static/html/ps3plot_btran1.html"
 }%

\edXgitlink{\giturl}{Source TeX}
\end{edXproblem}


\end{edXvertical}


\end{edXsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{edXchapter}
